### 2014-07-31 text/RoaM_proceedings/paper.tex
proceedings final

\documentclass{llncs}

\usepackage[style=authoryear-comp, % Citation marks as [Jef65]
            natbib=true,      % Natbib-style cite macros \citeauthor &c.
            hyperref=true,    % Cites in pdf are links to bib
                              % (hyperref conf.)
            maxnames=3,       % truncate name lists if more than 2
                              % names appear
            doi=false,        % no doi's
            url=false,        % no url's
            sortcites=false,  % do NOT sort cites in the style of the
                              % bibliography 
            %backref=true     % insert backrefs in reference section
            ]{biblatex}
\bibliography{MyRefGlobal}

\usepackage{amsfonts}
\usepackage{dsfont}

% \usepackage[colorinlistoftodos,color=lightgray,bordercolor=blue,textsize=footnotesize]{todonotes}

% \usepackage{setspace}
% \doublespacing
% \usepackage[left=3.75cm,right=3.75cm,top=3.75cm,bottom=3.75cm]{geometry}

\title{On the Evolution of Choice Principles}
\titlerunning{On the Evolution of Choice Principles}
\author{Michael Franke\inst{1} \and Paolo Galeazzi \inst{2}}
\institute{Department of Linguistics, University of T\"{u}bingen \and
  Institute for Logic, Language and Computation, Universiteit van Amsterdam}
\date{}

\begin{document}

\maketitle

\begin{abstract}
  We compare level-$k$ utility maximization and level-$k$ regret
  minimization in evolutionary competition based on averages of
  one-shot plays over many randomly generated strategic games. Under
  the assumption that Theory-of-Mind-reasoning of depth $k$ incurs a
  cost monotonically increasing with $k$, our results show that mixed
  states with low levels of reasoning depth $k$ can be evolutionary
  stable with substantial basins of attraction under the replicator
  dynamics and that regret minimization can outperform utility
  maximization.
\end{abstract}

\noindent The widely accepted normative standard of individual
decision making under uncertainty is expected utility maximization. An
agent is rational only if his choices maximize expected utility. In
games with several players, agents who believe that their opponents
are rational might thus rule out certain of the opponents'
choices. This kind of reasoning can be iterated, giving rise to
predictions of rational choice under arbitrarily deep nestings of
mutual beliefs in
rationality. % Depending on the detailed specification of what it means
% to have a belief in the opponents' rationality, and to respond
% rationally to such beliefs, standard game-theoretic solution concepts
% are characterized by this nested Theory of Mind reasoning, most
% notably the iterated deletion of strongly dominated strategies in
% strategic games \citep[e.g.][for
% overview]{Perea2012:Epistemic-Game-}.

Empirical data on human solitary or interactive decision-making is at
odds with this idealized picture. For one, human decision makers
systematically deviate from the predictions of expected utility theory
\citep[e.g.][]{TverskyKahnemann1974:Judgement-under,TverskyKahnemann1981:The-Framing-of-}. For
another, human capability for higher-order Theory of Mind (ToM)
reasoning appears to be rather limited \citep[e.g.][inter
alia]{HoCamerer1998:Iterated-Domina,KeyzarLin2003:Limits-on-Theor,VerbruggeMol2008:Learning-to-App,DegenFranke2013:Cost-Based-Prag}.

The usual reactions to this discrepancy between idealized norm and
empirical observation is to either rethink the prescriptive axioms of
rational choice or, alternatively, to devise a descriptive theory to
capture the empirical facts of human psychology of decision-making and
strategic reasoning
\citep[e.g.][]{Camerer2003:Behavioral-Game,GlimcherCamerer2009:Neuroeconomics:}. The
approach taken here is different from either. We refer to a general
decision making rule as a \emph{choice principle}. Formally, a choice
principle is a function that takes a game as input and returns a set
of acts as the decision maker's choice. We then adopt an evolutionary
point of view and ask: which choice principles are successful on
average in recurrent competition with alternative choice principles
(including more or less limited ToM reasoning)?

\citet{Weerdde-WeerdVerbrugge2013:How-much-does-i} address a part of
this question by looking at agent-based simulations of utility
maximizers with different ToM reasoning capabilities when playing
repeatedly selected zero-sum games against each other. The simulations
of \citeauthor{Weerdde-WeerdVerbrugge2013:How-much-does-i} suggest
that there is a benefit to deeper ToM reasoning only up to a fairly
limited depth of reasoning.

Extending and generalizing this line of investigation, we use
numerical simulations to approximate the expected payoff of agents
with a fixed choice principle who play many \emph{arbitrary} 2-player
one-shot strategic games, not just a small hand-picked selection. This
way we try to assess the general evolutionary benefit of a choice
principle across an unbiased sample of games, not just for those games
that are of particular technical, philosophical or historical
importance. Additionally, we compare not only different ToM reasoning
capability of utility maximizers, but also variation in the underlying
method of choice. Concretely, we look at level-$k$ utility
maximization and level-$k$ iterated regret minimization (IRM)
\citep{HalpernPass2012:Iterated-Regret}. The motivation for looking at
IRM is threefold. For one, regret minimization has been suggested as a
conceptually appealing alternative to utility maximization and a
potentially promising descriptive theory of individual decision-making
\citep{LoomesSugden1982:Regret-Theory:-}. For another, we are able to
show that non-iterated regret minimization outperforms the closely
related MaxiMin security strategy (see below) in evolutionary settings
like the one investigated here, making regret minimization the
evolutionarily best representative of a security strategy that we are
aware of.\footnote{More concretely, so far we have been able to prove
  this result for the class of 2-player symmetric games, and to
  demonstrate that it holds more generally using numerical simulations
  like the one reported here.} Finally, IRM usually requires only few
iteration steps, making it a serious challenger of the standard theory
in terms of limited ToM reasoning as well
\citep[c.f.][]{HalpernPass2012:Iterated-Regret}.

The choice principles we compare are defined as follows. Look at
level-$k$ utility maximizers first. L-0 utility maximizers have an
arbitrary probabilistic belief about the opponent's behavior and are
indifferent between any act that maximizes expected utility under this
belief. L-$k+1$ utility maximizers have an arbitrary probabilistic
belief with support only on acts that an L-$k$ opponent might choose,
and are indifferent between any act that maximizes expected utility
under this belief. Notice that this construction is different from
that of \citet{Weerdde-WeerdVerbrugge2013:How-much-does-i} in several
respects. Firstly, we look at level-$k$ reasoning models
\citep[e.g.][]{Crawford2003:Lying-for-Strat,Crawford2007:Lets-Talk-It-Ove},
not cognitive hierarchy models
\citep[e.g.][]{CamererHo2004:A-Cognitive-Hie}, like
\citeauthor{Weerdde-WeerdVerbrugge2013:How-much-does-i} do. The former
models assume that agents of level $k+1$ believe that the opponent is
of level $k$, while the latter have a more general belief that the
opponent is of some level-$l$ with $l \le k$. Moreover, we assume all
games to be one-shot encounters, while
\citeauthor{Weerdde-WeerdVerbrugge2013:How-much-does-i} look at
repeated encounters with the same opponent, and the concomitant
possibility of strategic learning.

Let's turn next to the definition of level-$k$ regret
minimizers. Formally, if $U$ is a matrix of utilities for the row
player, regret minimization is equivalent to playing MaxiMin on the
derived matrix $U'$, where $U'_{ij} = U_{ij} - \textrm{max}_{k}
U_{kj}$ is the negative regret of choosing act $i$ when the opponent
chooses $j$. L-0 regret minimizers are indifferent between any act
that minimizes regret in this sense. An L-$k+1$ regret minimizer
chooses like an L-0 regret minimizer in the reduced game in which only
acts remain that row and column players would choose that are L-$k$
regret minimizers \citep[see][]{HalpernPass2012:Iterated-Regret}.

We randomly generate arbitrary strategic games and record the payoff
earned by each choice principle when playing against any other,
including itself. In general, the details of generating arbitrary
games are important to this approach. Some choice principles might be
better in certain classes of games, but not others. For that reason,
we used several algorithms for creating games that are as unbiased and
neutral as possible. We choose neutrality in the selection of games,
because we have at present no better answer to the main empirical
question that is relevant here, namely which kinds of interactive
decision-making situations were most frequent during the critical
stages of cognitive development. For illustration, this abstract
focuses on the results of a simple generic algorithm for constructing
arbitrary asymmetric strategic games with 2 players. The procedure
starts by first picking a random number between 2 and 10 as the number
of acts for each player, say $n$ and $m$. Then, we determine each
entry in the $n \times m$ and $m \times n$ utility matrices of each
player as an integer between 0 and 10, all independently and uniformly
at random.

\begin{table}[t]
  \centering
\begin{tabular}{rrrrrrrrr}
  \hline
 & EU-L0 & EU-L1 & EU-L2 & EU-L3 & EU-L4 & RM-L0 & RM-L1 & RM-L2 \\ 
  \hline
  EU-L0 & 6.735 & 6.749 & 6.739 & 6.873 & 6.755 & 6.742 & 6.733 & 6.734 \\ 
  EU-L1 & 7.126 & 6.650 & 6.649 & 6.610 & 6.954 & 6.969 & 7.016 & 7.017 \\ 
  EU-L2 & 6.642 & 7.413 & 6.979 & 6.857 & 6.961 & 6.414 & 6.494 & 6.495 \\ 
  EU-L3 & 6.636 & 6.946 & 7.627 & 7.152 & 7.162 & 6.430 & 6.465 & 6.465 \\ 
  EU-L4 & 6.780 & 7.038 & 7.314 & 7.734 & 7.388 & 6.592 & 6.620 & 6.619 \\ 
  RM-L0 & 6.686 & 6.684 & 6.700 & 6.830 & 6.695 & 6.681 & 6.676 & 6.675 \\ 
  RM-L1 & 6.853 & 6.768 & 6.782 & 6.913 & 6.806 & 7.051 & 7.046 & 7.043 \\ 
  RM-L2 & 6.856 & 6.770 & 6.784 & 6.916 & 6.810 & 7.052 & 7.057 & 7.053 \\ 
   \hline \\
\end{tabular}
\caption{Average utilities of level-$k$ utility maximizers and
  level-$k$ regret minimizers over 10.000 arbitrary games (see main
  text). In our sample, regret minimizers of level $k>2$ are
  behaviorally equivalent to RM-L$2$ and are therefore omitted in this table. } 
  \label{tab:MaxUtil}
\end{table}

Table~\ref{tab:MaxUtil} gives the accumulated payoff averaged over
10.000 games sampled in this way, for $0 \le k \le 4$. The table lists
the average payoff for the row principle when playing against the
column principle and can be regarded as a symmetric ``meta-game'' that
captures the evolutionary competition between choice principles. The
only evolutionarily stable strategy (ESS) of this meta-game is EU-L4,
but it is clear (from more extensive simulations) that this is just
because we chose an arbitrary cutoff for depth of
reasoning. Generally, it appears that pure populations of EU-L$k$
players can always be invaded by EU-L$k+1$ players. But,
interestingly, some pure populations of EU-L$k$ players can also be
invaded by EU-L$l$ players, with $l<k$ (see
Table~\ref{tab:MaxUtil}). On the other hand, if $k\ge2$, then RM-L$k$
is a neutrally stable strategy (NSS): regret minimizers of a different
level $l\ge2$ would not be driven to extinction, but would also not
take over the population (because they are behaviorally equivalent in
all games from our random sample).

The results from this example generalize to other game-sampling
routines, as long as they are general and encompassing enough.  While
there is usually no $k$ at which EU-L$k$ is an ESS, there is a small
$k$ (almost always $k\le 3$) for which IRM has reached its fixed point
in all sampled games and RM-L$k'$ is an NSS for all $k' \ge k$. This
is a noteworthy result: on randomly sampled games, it is always
beneficial for utility maximizers to outsmart the opponent by ideally
one level of ToM reasoning, but too much ToM outsmarting is
harmful. For IRM, levels of ToM reasoning higher than a few steps are
evolutionarily pointless.

More fine-grained results obtain when we look at the predictions of
the replicator dynamics on the simulated ``meta-games'' together with
the natural hypothesis that ToM reasoning incurs a cost $c(k) \in
\mathds{R}$, which is monotonically increasing with $k$. For
concreteness, we look at a parameterized cost function $c(k \thinspace
; \thinspace s, a ) = \sum_{i=0}^{k} a^i \thinspace s$ where $s$ is
the initial step of growth, and $a$ is a parameter that controls for
superlinear and sublinear growth. For at least linear growth ($a \ge
1$), there is a threshold $\theta \approx .391$ such that if $s >
\theta$, the only attractor is EU-L$0$. With $s$ smaller or not too
much bigger than $\theta$ and sublinear growth (e.g., $a=0.25$), RM-L1
is a strong attractor, as well as mixed states of low level
EU-maximizers. For step $s < \theta$ and $a \le 1$ results are more varied. If $.313
< s < \theta$, the only attracting state is a mix of EU-L0 and
EU-L1. If $.011 < s < .313$, RM-L1 is a strict symmetric Nash
equilibrium and therefore an attractor, but also, depending on $s$,
mixed states of low-level EU-maximizers are attracting. If $0 < s \le
.011$ RM-L2 is an attractor, together with mixed states of low level
EU-maximizers.

In sum, our simulation data show that shallow depth of reasoning is
plausibly evolutionarily dominant in long run competition with ``deep
ToM'' strategies. Regret minimization often outperforms utility
maximization in this evolutionary competition, especially when
reasoning costs are added. Although already insightful, our results
are only partial, because many interesting questions have not yet been
addressed. We mention just three obvious extensions for future
work. Firstly, sequential games might yield different results, because
they would induce more ties in strategic form, and hence more reasons
to apply deeper ToM reasoning. Secondly, the payoff of utility
maximization heavily depends on the type of belief about the
opponent's choice. Here we used arbitrary beliefs compatible with the
opponent's assumed type, but certain restrictions on belief formation,
e.g., using maximum-entropy beliefs, might give utility maximizers a
better payoff on average. Finally, it would be interesting to extend
our comparison to more choice principles, e.g., ones including
learning from repeated interactions with the same players
\citep[e.g.][]{Weerdde-WeerdVerbrugge2013:How-much-does-i}, or choice
principles building on prospect theory, for example.

\newpage

\printbibliography[heading=bibintoc]

\end{document}