### 2015-12-14 paper/graph.tex
graph added

\documentclass{article}

\usepackage[]{mypackages}

\usepackage{mycommands}

\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{5pt}

\usetikzlibrary{positioning,arrows,calc,fit}

% \usepackage{pgfplots}

\begin{document}

\begin{tikzpicture}[node distance = 2.5cm]

  \begin{scope}

    \node[] (environment) {};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = environment, node distance = 2cm] (situation)
    {\textbf{decision situation} $d$};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = situation] (representation) 
    {\textbf{subjective representation} $s$ of $d$ \\ \textcolor{gray}{[preference, beliefs, \dots]}};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text centered, below of = representation] (action)
    {\textbf{action choice} $a$ based on $s$};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = action] (payoff) {\textbf{payoff} for $a$ in $d$};

    \draw[->, thick]
    ([xshift=-1.75cm]environment.center)--([xshift=-1.75cm,yshift=0.5cm]situation.center)
    node[midway, right] {environment presents random $d$};

    \draw[->, thick]
    ([xshift=-1.75cm,yshift=-0.5cm]situation.center)--([xshift=-1.75cm,yshift=0.55cm]representation.center)
    node[midway, right] {subj.~representation scheme $\gamma$};

    \draw[->, thick]
    ([xshift=-1.75cm,yshift=-0.55cm]representation.center)--([xshift=-1.75cm,yshift=0.5cm]action.center)
    node[midway, right] {action selection function $\kappa$};

    \draw[->, thick]
    ([xshift=-1.75cm,yshift=-0.55cm]action.center)--([xshift=-1.75cm,yshift=0.5cm]payoff.center)
    node[midway, right] {};

    \node[minimum height = 0.75cm, text width=6cm, text
    centered, right of = representation, xshift = 2cm, rotate = 90] (mechanism) {choice mechanism};

    \begin{pgfonlayer}{background}
             \node [rounded corners,
       very thick,draw=black!40,fit={($(situation.south)+(0,+30pt)$) 
                                 ($(mechanism.west)+(+5pt,0)$) 
                                 ($(action.north)+(0,-30pt)$) 
                                 ($(representation.west)+(-7pt,0)$)}] (box) {};
    \end{pgfonlayer}

  \end{scope}

\end{tikzpicture}

\begin{tikzpicture}[]

  \begin{scope}

    \node (tpic) at ( -5,0)  {\includegraphics[width=4cm]{s1_vx.png}};

    \node (cpic) at (0,0)    {\includegraphics[width=4cm]{s1_v2.png}};

    \node (dpic) at ( 5,0)   {\includegraphics[width=4cm]{s2_v1.png}};

    \node (tlabel) at (-5, -4) {\Large{\mystate{t}}};

    \node (clabel) at (  0,-4) {\Large{$\mystate{c}^*$}};

    \node (dlabel) at (5,-4)   {\Large{\mystate{d}}};

    \node (mainlabel) at (0, -5.3) {\Large{Possible Referents}};

    \begin{pgfonlayer}{background}
       \node [thick,draw=black!40,fit=(tpic) (dlabel) (mainlabel) (dpic)] {};
    \end{pgfonlayer}
  \end{scope}

\begin{scope}[xshift = 7.5cm, yshift = -2.8cm, x=1.05cm, y = 1.1cm]

    \node (tpic) at  (0,3)    {\includegraphics[height=1.4cm]{v2.png}};
    \node (d1pic) at (3,3)    {\includegraphics[height=1.4cm]{s1.png}};
    \node (cpic) at  (0,0)    {\includegraphics[height=1.4cm]{s2.png}};
    \node (d2pic) at (3,0)    {\includegraphics[height=1.4cm]{v1.png}};

    \node (tlabel) at  (0,2)  {\Large{\mymessg{t}}};
    \node (d1label) at (3,2)  {\Large{$\mymessg{c}^*$}};
    \node (clabel) at  (0,-1) {\Large{\mymessg{d1}}};
    \node (d2label) at (3,-1) {\Large{\mymessg{d2}}};

    \node (mainlabel) at (1.5, 4.3) {\Large{Message Options}};

    \begin{pgfonlayer}{background}
       \node [thick,draw=black!40,fit=(tpic) (d2label) (mainlabel) (d2pic)] {};
    \end{pgfonlayer}
  \end{scope}

\end{tikzpicture}

\end{document}